---
title: "DSC 200 Lab Project -Term 2232"
date: "`r Sys.Date()`"
output: github_document
---

**Student Name:Mohmmad Sultan**

**Student ID: 2210002586**


**Deadline:** 23:59 on Sunday, 19 May 2024

**Total Points:** 20


## Loading Packages

```{r load-packages, message = FALSE}

library(tidyverse)
library(openintro)
library(ggrepel)
```


## Tasks

`1. (2 points)
```{r}
nrow(seattlepets)

```
Task 1: How many pets are included in this dataset?




`2. (2 points)

```{r}
ncol(seattlepets)

```
Task 2: How many variables do we have for each pet?



`3. (2 points)
```{r}
seattlepets %>%
  count(species)

```
Task 3: What are the pet species in Seattle? How many pets are there for each species?


`4. (2 points)
```{r}
seattlepets %>%
  count(animal_name) %>%
  arrange(desc(n)) %>%
  head(10)

```
Task 4: What are the ten most common pet names?


`5. (2 points)
```{r}
library(dplyr)

# Retrieve and display all the 6 records for the species Pig sorted by pet names
pigs <- seattlepets %>%
  filter(species == "Pig") %>%
  arrange(animal_name)

# Display the result
print(pigs)

```




`6. (2 points)
```{r}
# Retrieve and display ONLY the pet name (animal_name) and primary_breed of the species Goat sorted by pet names
goats <- seattlepets %>%
  filter(species == "Goat") %>%
  select(animal_name, primary_breed) %>%
  arrange(animal_name)

# Display the result
print(goats)

```





`7. (2 points)
```{r}
# Concatenate the two columns animal_name and species into a single column named pet
# then display license_number and pet sorted by pet
pets <- seattlepets %>%
  mutate(pet = paste(animal_name, species, sep = "; ")) %>%
  select(license_number, pet) %>%
  arrange(pet)

# Display the result
print(pets)

```



`8. (2 points)
```{r}
library(ggplot2)

# Plot the counts of the species as bars
ggplot(seattlepets, aes(x = species)) +
  geom_bar() +
  labs(title = "Counts of Pet Species in Seattle", x = "Species", y = "Count")

```




`9. (2 points)
```{r}
top_10_names <- seattlepets %>% 
filter(animal_name %in% c( "Lucy"  , "Charlie" , "Luna" , "Bella" , "Max"    , 
                           "Daisy" , "Molly"   , "Jack" , "Lily"  , "Stella" ))
top_10_names
```
`a.	What does the above code chunk do?
Filter the Dataset: It filters the seattlepets dataset to include only rows where the animal_name column matches one of the specified names in the list: "Lucy", "Charlie", "Luna", "Bella", "Max", "Daisy", "Molly", "Jack", "Lily", and "Stella".
Assign the Result to top_10_names: The filtered dataset is then stored in a new variable called top_10_names.
Display the Result: It displays the contents of top_10_names to show the filtered data


`b. Plot the counts of the pet names (animal_name) in top_10_names
```{r}
ggplot(top_10_names, aes(x = animal_name, fill = species)) +
  geom_bar() +
  labs(title = "Top 10 Pet Names in Seattle Segmented by Species", 
       x = "Pet Name", 
       y = "Count") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


`10.  (2 points)

`The below code plots the proportion of dogs with a given name versus the proportion of cats with the same name. The 20 most common cat and dog names are displayed. The diagonal line on the plot is the x = y line; if a name appeared on this line, the name’s popularity would be exactly the same for dogs and cats.

```{r eval = TRUE, echo = FALSE, fig.fullwidth = TRUE}
# data prep
name_props <- seattlepets %>% 
  filter(
    !is.na(animal_name),
    species %in% c("Dog", "Cat")
    ) %>%
  group_by(species) %>% 
  count(animal_name, sort = TRUE) %>%
  mutate(prop = n / sum(n))
    
cat_name_props <- name_props %>%
  filter(species == "Cat") %>%
  rename(cat_prop = prop) %>%
  slice(1:30)

dog_name_props <- name_props %>%
  filter(species == "Dog") %>%
  rename(dog_prop = prop) %>%
  slice(1:30)
    
comb_name_props <- inner_join(cat_name_props, dog_name_props, 
                              by = "animal_name") %>%
  ungroup() %>%
  select(animal_name, cat_prop, dog_prop)
    
# create viz
ggplot(comb_name_props, aes(x = cat_prop, y = dog_prop)) +
  geom_abline(intercept = 0, 
              color = COL["lgray","full"], 
              alpha = 0.8, 
              size = 1.5) +
  geom_text_repel(aes(label = animal_name), 
                  segment.color = COL["gray","full"], 
                  seed = 291252, max.iter = 10000) +
  geom_point(color = COL["blue","full"], alpha = 0.8) +
  theme_minimal() +
  labs(x = "Proportion of cats", y = "Proportion of dogs") +
  xlim(0.002, 0.01) +
  ylim(0.002, 0.01) +
  ggimage::geom_emoji(
      image = "1f436", aes(x = 0.003, y = 0.009), size = 0.1
      ) +
  ggimage::geom_emoji(
      image = "1f431", aes(x = 0.009, y = 0.003), size = 0.1
      )
```

`What names are more common for cats than dogs? The ones above the line or the ones below the line?

`Answer here.The names that are more common for cats than dogs are the ones above the line. This is because the y-axis represents the proportion of dogs and the x-axis represents the proportion of cats. Points above the line 
y=x indicate that the proportion for cats is higher than for dogs.



`Is the relationship between the two variables (proportion of cats with a given name and proportion of dogs with a given name) positive or negative? What does this mean in context of the data?

`Answer here 
The relationship between the two variables (proportion of cats with a given name and proportion of dogs with a given name) is positive. This means that, generally, if a name is popular among cats, it also tends to be popular among dogs. The points are scattered around the line 

y=x, indicating that the same names are often used for both cats and dogs, with varying degrees of popularity.

